Article Title

### 2013-09-25 sections/Figure1.tex
ongoing edits

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{figures/Figure1.pdf}
\end{center}
\caption{
{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\tt A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens. {\tt B.} Two-photon fast 3D imaging of calcium signals in an awake mouse. {\tt C.} Deconvolved calcium signals. {\tt D.} The sample covariance matrix of neuronal calcium signals in 200 ms time bins. 
}
\label{Figure_label}
\end{figure}

>>>>

\begin{figure}[htp]
\begin{center}
\includegraphics[width=2.5in]{figures/Figure1.pdf}
\end{center}
\caption{
{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\bf A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens. {\bf B.} Two-photon fast 3D imaging of calcium signals in an awake mouse. {\bf C.} Deconvolved calcium signals. {\bf D.} The sample covariance matrix of neuronal calcium signals in 200 ms time bins. 
}
\label{Figure_label}
\end{figure}

### 2013-09-25 sections/Results.tex
ongoing edits

\subsection*{Data acquisition}
We recorded the calcium activity  of dense populations of neurons in the supragranular layers of mouse primary visual cortex using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}   while presenting numerious repetitions of full-field drifting gratings (Fig. 1A and 1B). This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4. Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig. 1C and 1D).  The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig. 1E).

@@

\subsection*{Covariance estimation}
We aim to estimate the true covariance matrix $\Sigma = \mathbb E [(x-\mu)(x-\mu)^\T]$ , where $\mathbb E[\cdot]$ denotes expectation \TODO{under the true model}, $x$ denotes the $p\times 1$ vector of real-valued instanteous firing rates of $p$ neurons in a bin of duration $\Delta t$ and $\mu = \mathbb E[x]$.  For more rigourous notation, definitions, and derivations, see Appendix.

@@

\subsubsection*{Regularized estimates}
Although $\hat\Sigma_0$ is unbiased, it is not as close as possible to $\Sigma$.  Closer estimates can be produced by admitting a moderate amount of bias (\emph{approximation error}) in order to reduce the variance (\emph{estimation error}) of the estimate. 
\emph{Regularization} is the deliberate biasing of the estimate toward a low-dimensional target estimate in order to strike a favorable balance between bias and variance and minimize the overall expected error \cite{Bickel:2006,Ledoit:2004}.

A great variety of regularized covariance estimators can be conceived, which consistently outperform the sample covariance estimate. Regularized estimators differ from each other by the following aspects:
\begin{itemize}
\item the choice of the family of the target estimate, 
\item the choice of the specific member of the family of target estimates
\item the choice of the path from the unbiased estimate toward the chosen target
\item the distance along that path
\end{itemize}
Steps 1 and 2 produce a low-dimensional estimate. Steps 3 and 4 are often called shrinkage and allow deviations from (or distribution about) the low-dimensional estimates.

We evaluated four regularized estimators which we will denote as A, B, C, and D.  Their target estimates are graphically depicted in Figure 2. 
The green spheres represent the observed neurons. The light-colored balls represent latent units and the edges connecting them represent conditional dependencies.

\subsubsection*{Estimate A: Shrinkage toward the independent model}
In A, the target is a family of diagonal matrices 
\begin{equation}
\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I
\end{equation}
where $\circ$ is the entrywise product (Hadamard product). When $\alpha=0$, $\hat T_\alpha$ has the average variance along the diagonal. When $ \alpha=1$, $ \hat T_\alpha$ has empirical variances on the diagonal.

@@

\subsubsection*{Estimate C}
In estimate C, the target $ \hat S_\alpha$ is a matrix whose inverse is <em>sparse</em>, with a large fraction of off-diagonal elements fixed at zero. The coefficient $ \alpha$ controls the sparsity in $ \hat S_\alpha$.

>>>>

\subsection*{Data acquisition}
We recorded the calcium activity  of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.  Numerious repetitions of full-field drifting gratings (Fig. 1A and 1B) were presented to the eye contralateral to the imaged site. This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4. Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig. 1C and 1D).  The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig. 1E).

@@

\subsection*{Sample covariance matrix}
We aim to estimate the true covariance matrix $\Sigma = \mathbb E [(x-\mu)(x-\mu)^\T]$ , where $\mathbb E[\cdot]$ denotes expectation \TODO{under the true model}, $x$ denotes the $p\times 1$ vector of real-valued instanteous firing rates of $p$ neurons in a bin of duration $\Delta t$ and $\mu = \mathbb E[x]$.  For more rigourous notation, definitions, and derivations, see Appendix.

@@

\subsection*{Regularized estimates of the covariance matrix}
To assess the quality of a covariance matrix estimate $\hat\Sigma$, we must define a real-valued \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which attains its minimum when $\hat\Sigma = \Sigma$. For the purposes of this study, we adopted the \emph{Gaussian loss} function:
\begin{equation}
\mathcal L(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) = \frac 1 p(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma))
\end{equation}
which resembles the negative multivariate normal log-likelihood (see Appendix).

Although the sample covariance matrix is unbiased: $\mathbb E[\hat\Sigma_0]=\Sigma$, it is not as close to $\Sigma$ as possible, on average. We must seek  to minimize the expected value of the loss function, known as \emph{estimator risk} 
\begin{equation}
r = \mathbb E[\mathcal L(\hat\Sigma, \Sigma)]
\end{equation}

Estimator risk can be reduced by \emph{regularization}. Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.  Resulting regularized estimators are biased but  less variable. Regularization procedures must strike favorable balance between bias and variance so that the estimator risk is minimized. The terms \emph{bias} and \emph{variance} are applicable only the loss function is mean-squared-error. The more general terms are \emph{approximation error} and \emph{estimation error})

Numerous regularized covariance matrix estimates have been devised. They start with the sample covariance matrix but differ in the choice of the target estimate and the shrinkage path and intensity. Some regularization schemes focus on the optimal choice of the low-dimensional target estimate from a large family of target estimates and do not require shrinkages.  Others have a fairly simple target estimate but instead adjust the shrinkage intensity to reduce the estimation error.

A regularized estimate of $\Sigma$ can be constructed by either (a) selecting an optimal low-dimensional approximation of $\hat\Sigma_0$ from a given family of approximations and/or (b) biasing or \emph{shrinkage} of $\hat\Sigma_0$ toward a given low-dimensional target estimate by an optimal amount.  Both dimensionality reduction and shrinkage approaches reduce the variance of the solution while introducing bias.

We evaluated four regularized estimators which we will denote as A, B, C, and D.  Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2.  Here the green spheres represent the recorded neurons. The light-colored balls represent latent units of the graphical model and the edges connecting them represent conditional dependencies.

\input{sections/Figure2.tex}

A \emph{graphical model} is a multivarate probability distribution with a specified graph of conditional dependencies between its variables \cite{Koller:2009}.  When such a distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} or, equivalently, a \emph{Gaussian Markov Random Field}.  Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.

Fitting a graphical model to data involves two tasks: (1) the selection of the set of non-zero covariances or \emph{covariance selection} \cite{Dempster:1972} and (b) the fitting of the non-zero elements.

\subsubsection*{Estimate A: Shrinkage toward the independent model}
In estimate A, the target is a family of diagonal matrices 
\begin{equation}
\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I
\end{equation}
where $\circ$ is the entrywise product (Hadamard product). When $\alpha=0$, $\hat T_\alpha$ has the average variance along the diagonal. When $ \alpha=1$, $ \hat T_\alpha$ has empirical variances on the diagonal.

@@

\subsubsection*{Estimate C}
In estimate C, the target $ \hat S_\alpha$ is a matrix whose inverse is \emph{sparse}, with a large fraction of off-diagonal elements fixed at zero. The coefficient $ \alpha$ controls the sparsity in $ \hat S_\alpha$.

### 2013-09-25 sections/Theory.tex
ongoing edits

The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions. When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{t=1}^n x(t) x(t)^\T$ is  
\begin{equation}
L\left(\Sigma \mid \Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
\end{equation}
Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
\begin{equation}\label{eq:GaussLoss}
\mathcal L_g(\hat\Sigma,\Sigma) 
=  -\frac 2 {pn} L\left(\hat\Sigma \mid \Sigma \right) - \frac 1 p \ln(2\pi) 
\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma) \right) 
\end{equation}
\TODO{Note that the switch from $L\left(\Sigma \mid \hat\Sigma \right)$ to $L\left(\hat\Sigma \mid \Sigma \right)$ is on purpose, which will become clear in the next few steps.}
The corresponding excess loss 
\begin{equation}
\ell_g(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) - \mathcal L_g(\Sigma,\Sigma)  
= \frac 1 p \left(-\ln \det (\hat \Sigma^{-1} \Sigma) + \Tr(\hat \Sigma^{-1}\Sigma)\right) - 1
\end{equation}
is known as \emph{entropy loss} \cite{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}

@@

\subsection*{Regularization}

\emph{Regularization} is the deliberate biasing (`shrinkage') of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to minimize the estimator risk by striking a favorable balance between bias and variance.
A regularized covariance matrix estimator must solve two problems: (a) choose and fit the target covariate estimate $\hat T$ and (b) mix $\hat\Sigma_0$ with $\hat T$ by the optimal amount. 
Regularized covariance matrix estimates can be generally expressed as 
\begin{equation}
\hat\Sigma_{d,\lambda} = mix(\Sigma_0,\hat T_{\hat d},\hat\lambda) 
\end{equation}
where $\hat d$ indicates the choice of the target estimate from the family of target estimates $\hat T_d$. $mix(\cdot,\cdot,\hat\lambda)$ is the mixing function of $\hat\Sigma_0$ and $\hat T_{\hat d}$ in proportion controlled by $\hat \lambda$. Both $\hat d$ and $\hat \lambda$ must be calculated from the training sample.

A number of regularization schemes have been developed, which differ by their target estimates and mixing functions.    Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of $\hat\Sigma_0$.  The perplexing phenomenon that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \cite{Efron:1977}.   \TODO{This point can be untuitive.  For example, \cite{Varoquaux:2012} misrepresents the effect of regularization by stating that regularization removes the upward bias of correlations in the sample covariance matrix. In reality, regularization should start with an unbiased estimator and bias it toward a low-variance target so that the estimation risk is reduced.} However, if a family of target estimates can capture the underlying regularities in the covariance structure of a specific system with a small number of parameters, biasing the estimate toward these targets ought to produce a greater reduction in estimator risk.

In this study, we evaluate four regularized estimators whose target estimates correspond to the graphical models in Figure \ref{fig:02}: diagonal (A), multifactor (B), sparse partial correlations (C), and sparse  partial correlations with latent units (D).  The following describe each regularization scheme separately.

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure2.pdf}
\caption{
Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent. 
\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units. 
\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes 
}\label{fig:02}
\end{figure}

>>>>

The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions. When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{t=1}^n x(t) x(t)^\T$ is  
\begin{equation}
L\left(\Sigma \mid \hat\Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
\end{equation}
Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
\begin{equation}\label{eq:GaussLoss}
\mathcal L_g(\hat\Sigma,\Sigma) 
=  -\frac 2 {pn} L\left(\hat\Sigma \mid \Sigma \right) - \frac 1 p \ln(2\pi) 
\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma) \right) 
\end{equation}
\TODO{Note that the switch from $L\left(\Sigma \mid \hat\Sigma \right)$ to $L\left(\hat\Sigma \mid \Sigma \right)$ is on purpose, which will become clear in the next few steps.}
The corresponding excess loss 
\begin{equation}
\ell_g(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) - \mathcal L_g(\Sigma,\Sigma)  
= \frac 1 p \left(-\ln \det (\hat \Sigma^{-1} \Sigma) + \Tr(\hat \Sigma^{-1}\Sigma)\right) - 1
\end{equation}
is known as \emph{entropy loss} \cite{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}

@@
